---
title: "Data Cleaning"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)

```
# Read In the Data


```{r}
catch_original<-read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method="libcurl"), stringsAsFactors=FALSE)

head(catch_original)
```
# Tidy Up Data

Remove the notes and 'All' column because we don't need them

```{r}
catch_clean<-catch_original %>% 
  select(-All,-notesRegCode)%>%
  mutate(Chinook = ifelse(Chinook=="I",1,(Chinook))) %>% 
     mutate(Chinook = as.numeric(Chinook))      
  
head(catch_clean)
```


## Data QA

DEtermine which values were converted to NA by as.numeric
Checking for NA values


```{r}
i<- which(is.na(catch_clean$Chinook))
catch_original[i,]
```
 explain what was done with I in row 401

## Convert to long form

Data table iis in wide format and we want long

```{r}
catch_long<-catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to ="Catch") %>% 
  rename(catch_thousands=Catch) %>% 
  mutate(Catch = catch_thousands*1000) %>% 
  select(-catch_thousands)

```


# Summarize

CAlculate the mean catch by region

```{r}
mean_region<-catch_long %>% 
  group_by(Region, species) %>% 
  summarize(mean_catch=mean(Catch),n_obs=n())
head(mean_region)
```

Filter for only sockeye





# Join to other Data
```

```{r}
regions_defs<-read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method="libcurl"), stringsAsFactors=FALSE) %>% 
  select(code,mgmtArea)
head(regions_defs)
```

```{r}
catch_joined<-left_join(catch_long, regions_defs, by = c("Region" ="code"))
head(catch_joined)
```






```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAI-039"),
                       stringsAsFactors = FALSE) %>% 


  separate(site, c("island", "site_number"), sep = "-")
head(sites_df)
```
```{r}
dates_df <- data.frame(year = c("1930",
                                "1930",
                                "1930"),
                       month = c("12",
                                "12",
                                "12"),
                       day = c("14",
                               "15",
                               "16"),
                       stringsAsFactors = FALSE) %>% 
  unite(date, year, month, day, sep = "-")
head(dates_df)
```



